Type-Safe Modular Parsing

### 2016-12-20 paper/sections/Overview.tex
Summary, rewrite one example

\bruno{Write a summary for this section}
\bruno{Fill in all the missing references}

@@

\huang{done, the example may be not very good, but i haven't found a
  better one}
\bruno{This example does not appear to be problematic: adding ``-''
  does not affect the rest of the grammar. Isn't the problem that
  sometimes adding a new case to the grammar, requires changing other
parts? Otherwise I don't see why left-recursion is a problem, and we
cannot simply claim that it is a problem. You have to make a better
effort to find a problematic grammar.
}

\paragraph{Backtracking} The need for backtracking is also problematic
in a modular setting. Take Parsec as an example, its choice
combinator only tries the second alternative if the first fails
without any token consumption. If two alternatives have a same
non-empty prefix, an auxiliary function \inlinecode{try} must be added
to backtrack. However, if we want to extend our parsers later, we
should always consider the worst case in which all alternatives share
common prefixes. Then we need to add \inlinecode{try} for all the
branches, which results worst case exponential time complexity.

For example, suppose that we only have \inlinecode{import..from} statements before, now we want to extend the grammar to add an \inlinecode{import..as} case. The third line of grammar below is the new case.

@@

Because the new case \inlinecode{import..as} shares a prefix with the old case \inlinecode{import..from}, we must backtrack in order to cover all possibilities.

\begin{lstlisting}[language=PlainCode]
oldParser = parseImportFrom <|> ...
newParser = try parseImportFrom <|> ... <|> parseImportAs
\end{lstlisting}

\bruno{The example is good, I think but the explanation is not.
You want to say that when you know
  the full grammar, you can figure out where ``try'' is
  needed. Without the full grammar you'd need assume the worst
  case. Make an effort to make your explanation cristal clear!
I think you want to miss the abstract explanation that you give first,
with the explanation about the concrete example.
Start with ``The need for backtracking is also problematic
in a modular setting. For example, suppose ...'' and synchronize
the abstract explanation and the explanation for the example.
}

@@

\subsection{The Parsing Expression Problem}
\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
  the solution without pointing out what the problem is first. What
  you need to do is: First show what happens with conventional
  parsers: at some point, if you add extensions the recursive calls
  will be wrong. Then you show (in the next section) the solution:
 use delegation/open recursion.}

@@

\subsection{Delegation-Based Parsing}\label{subsec:overview-firstlook}

>>>>

This section gives an overview of our library \name, from a user's perspective. Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions. We start from section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers. Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse. Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls. In section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers. Finally, section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.

\huang{done} \bruno{Write a summary for this section}

\huang{done}\bruno{Fill in all the missing references}

@@

\huang{todo}
\bruno{This example does not appear to be problematic: adding ``-''
  does not affect the rest of the grammar. Isn't the problem that
  sometimes adding a new case to the grammar, requires changing other
parts? Otherwise I don't see why left-recursion is a problem, and we
cannot simply claim that it is a problem. You have to make a better
effort to find a problematic grammar.
}

\paragraph{Backtracking} The need for backtracking is also problematic
in a modular setting. For example, suppose that we only have \inlinecode{import..from} statements before, now we want to extend the grammar to add an \inlinecode{import..as} case. The third line of grammar below is the new case.

@@

Because the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning. Take Parsec as an example, its choice
combinator \inlinecode{<|>} only tries the second alternative if the first fails
without any token consumption. An auxiliary function \inlinecode{try} is used for explicit backtracking.

\begin{lstlisting}[language=PlainCode]
oldParser = parseImpFrom <|> parseA <|> parseB <|> ...
newParser = try parseImpFrom <|> parseA <|> parseB <|> ... <|> parseImpAs
\end{lstlisting}

Given the full grammar, we can decide where to put \inlinecode{try} for backtracking. However, in a modular setting we should always consider the worst case, that all alternatives may share common prefixes with future cases. Then we need to backtrack for all the branches, which results worst case exponential time complexity.
In Parsec, we must add \inlinecode{try} everywhere:

\begin{lstlisting}[language=PlainCode]
parser = try parseImpFrom <|> try parseA <|> try parseB <|> ... <|> try parseImpAs
\end{lstlisting}

\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
You want to say that when you know
  the full grammar, you can figure out where ``try'' is
  needed. Without the full grammar you'd need assume the worst
  case. Make an effort to make your explanation cristal clear!
I think you want to miss the abstract explanation that you give first,
with the explanation about the concrete example.
Start with ``The need for backtracking is also problematic
in a modular setting. For example, suppose ...'' and synchronize
the abstract explanation and the explanation for the example.
}

@@

\subsection{The Parsing Expression Problem}\label{subsec:overview-problem}
\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
  the solution without pointing out what the problem is first. What
  you need to do is: First show what happens with conventional
  parsers: at some point, if you add extensions the recursive calls
  will be wrong. Then you show (in the next section) the solution:
 use delegation/open recursion.}

@@

\subsection{Delegation-Based Parsing}\label{subsec:overview-delegation}
